Generalized Bradley-Terry models

### 2025-03-04 inst/jss_hyper3_real.Rnw
more work on counterstrike

{\em Counter Strike: Global Offensive (CS:GO)} is a multiplayer
first-person shooter game in which two teams of five compete in an
immersive virtual reality combat environment.  Gameplay is typically
complex, including features such as cooperative teamwork, within-team
communication, and real-time strategic fluidity.

Here I analyse a dataset taken from recent esports tournament {\em
  ELEAGUE Major: Boston 2018}.  This Valve-sponsored Major featured 24
elite teams competing for a million dollar prize pool; peak concurrent
viewership exceeded $1.3\times 10^6$, the highest in CS:GO history at
the time.  I study the performance of tournament champions {\em
  Cloud9} who, as underdogs, narrowly defeated aggressive European
rivals {\em FaZe Clan} in the final.

The dataset used is the order statistic for number of deaths by each
of the five {\em Cloud9} players.  Deaths are

>>>>

\newcommand{\skad}{{\sc Skadoodle}}
\newcommand{\rush}{{\sc RUSH}}
\newcommand{\tarik}{{\sc tarik}}
\newcommand{\stew}{{\sc Stewie2K}}

{\em Counter Strike: Global Offensive (CS:GO)} is a multiplayer
first-person shooter game in which two teams of five compete in an
immersive virtual reality combat environment.  Gameplay is typically
complex, including features such as cooperative teamwork and real-time
strategic fluidity.

Here I analyse a dataset taken from recent e-sports tournament {\em
  ELEAGUE Major: Boston 2018}.  This Valve-sponsored Major featured 24
elite teams competing for a million dollar prize pool; peak concurrent
viewership exceeded $1.3\times 10^6$, the highest in CS:GO history at
the time.  I study the performance of tournament champions {\em
  Cloud9} who, as underdogs, narrowly defeated aggressive European
rivals {\em FaZe Clan} in the final.

Typically when a player is killed, the team suffers a number of
strategic disadvantages; player deaths are thus a quantifiable proxy
for both individual ability and also both teams' tactical execution.
Here I analyse the order statistic for number of deaths of each of the
five {\em Cloud9} players.

<<echo=FALSE>>=
a <- read.table("cloud9_boston_2018.txt", header=TRUE)
a[,-1]
@

From the first line we see that {\em Cloud9} played {\em FaZe Clan} on
the {\em inferno} map, won the match, and \skad\ had the least number
of deaths, followed by \rush, then \tarik, and so on.  Consider two
{\em Cloud9} players: \skad\ and \stew.  These teammates, although
comparable in skill [we fail to reject $H_0\colon p_{\mbox{\tiny\sc
      Skadoodle}}\neq p_{\mbox{\tiny\sc Stewie2k}}$, Plackett-Luce,
  $p=0.28$] have playing styles that stand in dramatic contrast.
Commentators characterise \skad\ as emphasising reactive map control
while \stew\ is more of a disruptive presence engaging in
high-variance play.

Given that these two players offer equally attractive and accessible
targets to the opposing team, it is possible that some tactical
situations might result in a systematic preference for targetting
\skad, and some a preference for killing \stew.  If this is so, the
deaths will be overdispersed and one would have a red bus-blue bus
situation, but with $\lambda<1$.

To assess this, we follow exactly the same protocol as for the sushi
preference dataset but with modified Plackett-Luce likelihoods using
the deathorder statistic from the dataset above, and equivalence
classes comprising \skad\ and \stew, and everybody else:

<<>>=
ec <-c(autimatic=1, RUSH=1, Skadoodle=2, Stewie2K=2, tarik=1)
@

<<defredbusfunc, echo=FALSE>>=
mk <- function(lam){
  H3 <- hyper3()
  for(i in seq_len(nrow(a))){

    d <- strsplit(a$deathorder[i],", ")[[1]]
    H3 <- H3 + cheering3(d, e=ec, help=c(1,lam))
  }
  return(H3)
}
f <- function(l){maxp(mk(l), give=1)$`Log-likelihood`}
@

Find the MLE for $\lambda$:

<<findopt, cache=TRUE, echo=FALSE>>=
optlam <- optimize(f,c(0.09, 0.14), maximum = TRUE)
evaluate <- optlam$maximum
support <- optlam$objective - f(1)
pvalue <- pchisq(support,df=1,lower.tail=FALSE)
@

<<calcklikelam, cache=TRUE, echo=FALSE>>=
lam <- exp(seq(from = -4,to = log(1.1), len = 23))
L <- sapply(lam, f)
@

<<findinterval, cache=TRUE, echo=FALSE>>=
g <- function(l){f(l) - optlam$objective  + 2}
jj1 <- uniroot(g, interval = exp(c(-4,-3)))
jj2 <- uniroot(g, interval = exp(c(-1,-0)))
lowerlimit <- jj1$root
upperlimit <- jj2$root
@

\begin{figure}[htbp]
  \begin{center}
<<plotlikelam, echo=FALSE>>=
plot(log(lam), L - max(L), type='b')
segments(x0=log(optlam$maximum),y0 = -0.5, y1 = 0.1,col='red')
abline(h=c(0,-2), col = 'gray')
segments(x0 = log(jj1$root), y0 = -2.2, y1 = -1.8, col = 'blue')
segments(x0 = log(jj2$root), y0 = -2.2, y1 = -1.8, col = 'blue')
abline(v = 0, lty = 2)
@ 
\caption{A\label{profsupcounterstrike} profile support function for
  the red bus-blue bus parameter $\log\lambda$ applied to mutually
  supporting equivalence class
  $\left\lbrace\mbox{\skad},\mbox{\stew}\right\rbrace$.  Vertical gray
  line shows $H_0\colon\lambda =1$, for which the support is about
  $\Sexpr{round(support,2)}$ relative to the evaluate
  $\hat{\lambda}\simeq\Sexpr{round(evaluate,3)}$ (small red bar),
  corresponding to a $p$-value of about $\Sexpr{round(pvalue,3)}$.
  Horizontal lines show the two-units-of-support credible interval of
  $(\Sexpr{round(lowerlimit,3)},\Sexpr{round(upperlimit,3)})$, marked
  in blue}
\end{center}
\end{figure}
